---
title: "Working with Penguin Dataset"
author: "Ruth Kingsley"
date: "2/23/2022"
---


## Setting up my R environment by installing and loading the required packages
## Installing packages with install.packages()
```{r}
install.packages("tidyverse")
install.packages("lubridate")
install.packages("skimr")
install.packages("janitor")
install.packages("palmerpenguins")
```

### Loading Packages with library function
```{r}
library(tidyverse)
library(lubridate)
library(skimr)
library(janitor)
library(palmerpenguins)
```

## show summary of dataset
```{r}
skim_without_charts(penguins)
```

## View penguins data
```{r}
View(penguins)
```

To view the first 6 rows of the entire dataset, I use the head()
## look at the data
```{r}
head(penguins)
```

## View dataset structure and column names
```{r}
str(penguins)
```
## look at the column names
```{r}
colnames(penguins)
```
## Cleanin THE Penguin data
### changing the character case of column names to lower case
```{r}
rename_with(penguins, tolower)
```
## Check if the name formats are valid
```{r}
clean_names(penguins)
```

## Sorting Data with arrange() function
## The first example sorts in ascending order while the second sorts in descending order
```{r}
penguins %>%
  arrange(bill_length_mm)

penguins %>%
  arrange(-bill_length_mm)

penguins2 <- penguins %>% arrange(-bill_length_mm)
View(penguins2)
```
## Organizing penguin data using different functions in R
```{r}
penguins %>% group_by(island) %>% drop_na() %>% summarize(mean_bill_length_mm = 
                                                             mean(bill_length_mm))

penguins %>% group_by(island) %>% drop_na() %>% summarize(max_bill_length_mm = 
                                                            max(bill_length_mm))

penguins %>% group_by(species, island) %>% drop_na() %>% 
  summarize(max_bl = max(bill_length_mm), mean_bl = mean(bill_length_mm))

penguins %>% filter(species == "Adelie")

penguins %>% filter(bill_length_mm == 40)
```

## Creating Visuals for the Penguins Dataset
## Creating scatter plots for penguin data
```{r}
ggplot(data = penguins)+ geom_point(mapping = aes(x= flipper_length_mm, y= body_mass_g))

ggplot(data = penguins) + geom_bar(mapping = aes(x=species, fill= island))

ggplot(data = penguins) + geom_bar(mapping = aes(x=species, fill= island))+
  facet_wrap(~island)
```

## Creating annotation and labels for the visuals
```{r}
ggplot(data = penguins) +
  geom_smooth(mapping = aes(x= flipper_length_mm, y= body_mass_g)) +
  geom_point(mapping = aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  labs(title="Palmer Penguins: Body Mass vs. Flipper Length", subtitle ="Sample of Three Penguin Species",
       caption = "Data Collected by Kristen Gorman") +
  annotate("text", x=220, y = 3500, label="The Gentoos are the largest", color="purple",
           fontface="bold", size= 3)
```
