---
title: "Working with Penguin Dataset"
author: "Ruth Kingsley"
date: "2/23/2022"
output: html_document
---

## Setting up my Environment
Notes: Setting up my R environment by installing and loading the 'tidyverse' and 'palmerpenguins' packages

Installing packages with install.packages()

```{r installing packages}
install.packages("tidyverse")
install.packages("palmerpenguins")
```

Loading Packages with library function
```{r loading packages}
library(tidyverse)
library(palmerpenguins)
library(ggplot2)
```

To return a view of the penguins data, use the View(). Remember that **R is case sensitive**
``{r View penguins data}
View(penguins)
``

To view the first 6 rows of the entire dataset, I use the head()
```{r look at the data}
head(penguins)
```

The structure of this dataset and the column names can be viewed by the the following functions;
```{r View dataset structure and column names}
str(penguins)
```

```{r look at the column names}
colnames(penguins)
```

## Creating Visuals for the Penguins Dataset

To plot the visuals for the penguin data
```{r ggplot for penguin data visualization}
ggplot(data = penguins)+ geom_point(mapping = aes(x= flipper_length_mm, y= body_mass_g))
```

```{r}
ggplot(data = penguins) + geom_bar(mapping = aes(x=species, fill= island))
```

```{r}
ggplot(data = penguins) + geom_bar(mapping = aes(x=species, fill= island))+
  facet_wrap(~island)
```

Creating annotation and labels for the visuals
```{r warning=TRUE}
ggplot(data = penguins) +
  geom_smooth(mapping = aes(x= flipper_length_mm, y= body_mass_g)) +
  geom_point(mapping = aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  labs(title="Palmer Penguins: Body Mass vs. Flipper Length", subtitle ="Sample of Three Penguin Species",
       caption = "Data Collected by Kristen Gorman") +
  annotate("text", x=220, y = 3500, label="The Gentoos are the largest", color="purple",
           fontface="bold", size= 3)
```
