## ---
title: "Working with Penguin Dataset"
author: "Ruth Kingsley"
date: "2/23/2022"
## ---


## Setting up my R environment by installing and loading the required packages

## Installing packages with install.packages()

install.packages("tidyverse")
install.packages("lubridate")
install.packages("skimr")
install.packages("janitor")
install.packages("palmerpenguins")

### Loading Packages with library function
library(tidyverse)
library(lubridate)
library(skimr)
library(janitor)
library(palmerpenguins)

## show summary of dataset
skim_without_charts(penguins)

## View penguins data
View(penguins)


To view the first 6 rows of the entire dataset, I use the head()
## look at the data
head(penguins)

## View dataset structure and column names
str(penguins)

## look at the column names
colnames(penguins)

## Cleanin THE Penguin data
### changing the character case of column names to lower case
rename_with(penguins, tolower)

## Check if the name formats are valid
clean_names(penguins)

## Sorting Data with arrange() function
## The first example sorts in ascending order while the second sorts in descending order

penguins %>%
  arrange(bill_length_mm)

penguins %>%
  arrange(-bill_length_mm)

penguins2 <- penguins %>% arrange(-bill_length_mm)
View(penguins2)

## Organizing penguin data using different functions in R
penguins %>% group_by(island) %>% drop_na() %>% summarize(mean_bill_length_mm = 
                                                             mean(bill_length_mm))

penguins %>% group_by(island) %>% drop_na() %>% summarize(max_bill_length_mm = 
                                                            max(bill_length_mm))

penguins %>% group_by(species, island) %>% drop_na() %>% 
  summarize(max_bl = max(bill_length_mm), mean_bl = mean(bill_length_mm))

penguins %>% filter(species == "Adelie")

penguins %>% filter(bill_length_mm == 40)

## Creating Visuals for the Penguins Dataset
## Creating scatter plots for penguin data
ggplot(data = penguins)+ geom_point(mapping = aes(x= flipper_length_mm, y= body_mass_g))

```{r}
ggplot(data = penguins) + geom_bar(mapping = aes(x=species, fill= island))
```

```{r}
ggplot(data = penguins) + geom_bar(mapping = aes(x=species, fill= island))+
  facet_wrap(~island)
```

## Creating annotation and labels for the visuals
ggplot(data = penguins) +
  geom_smooth(mapping = aes(x= flipper_length_mm, y= body_mass_g)) +
  geom_point(mapping = aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  labs(title="Palmer Penguins: Body Mass vs. Flipper Length", subtitle ="Sample of Three Penguin Species",
       caption = "Data Collected by Kristen Gorman") +
  annotate("text", x=220, y = 3500, label="The Gentoos are the largest", color="purple",
           fontface="bold", size= 3)
